Distinguishing Spin-Aligned and Isotropic Black Hole Populations With Gravitational Waves

### 2017-03-28 paper/AlignedIsotropicGW.tex
Cole ack. IAS visiting fellows program.
Added appendix on rough, order-of-magnitude calculation.

We thank Richard O’Shaughnessy and Christopher Berry for discussions
comments on this work.

@@

>>>>

We thank Richard O’Shaughnessy and Christopher Berry for discussions
comments on this work.  MCM acknowledges support of the University of
Birmingham Institute for Advanced Study Distinguished Visiting Fellows
program.

@@

\subsection{Order of Magnitude Calculation}
\label{sec:om-odds-ratio}

It is possible to estimate at an order-of-magnitude level the rate at
which evidence accumulates in favour of or against the isotropic
models as more systems are detected.  Based on Figure
\ref{fig:chieff-distribution-models}, approximate the isotropic
population $\chieff$ distribution as uniform on
$\chieff \in \left[ -0.25, 0.25 \right]$ and the aligned population
$\chieff$ distribution as uniform on
$\chieff \in \left[0, 0.5\right]$.  Then the odds ratio between the
isotropic and aligned models for each event is approximately
\begin{equation}
  \label{eq:approx-odds}
  \frac{p\left( d \mid I \right)}{p\left( d \mid A \right)} \simeq
  \frac{P\left( -0.25 \leq \chieff \leq 0.25 \right)}{P\left( 0 \leq \chieff \leq 0.5 \right) },
\end{equation}
where $P\left( A \leq \chieff \leq B \right)$ is the posterior
probability that $\chieff$ is between $A$ and $B$.  Using our
approximations to the $\chieff$ posteriors described above, this gives
an odds ratio of $5$ in favour of the isotropic models, which is about
a factor of two smaller than the ratio in the more careful calculation
described in Section \ref{sec:O1}.  This is a satisfactory answer at
an order-of-magnitude level.

If the true distribution is isotropic and follows this simple model,
and our measurement uncertainties on $\chieff$ are $\simeq 0.2$, then
each subsequent measurement contributes on average a factor of
$\sim 2.5$ to the overall odds.  (There is a 1/3 chance each of the
critical value of $\chieff = 0$ lying at $-1\sigma$, $0$, and
$+1\sigma$ with respect to the centre of the posterior.)  After ten
additional events, then, the odds ratio becomes
$5 \times 2.5^{10} \simeq 5 \times 10^{5}$, or $4.3 \sigma$,
consistent with the results of the more detailed calculation described
above.
